TODO: A Survey of StarCraft AI Techniques

### 2012-05-20 survey.tex
Added Florian to acknowledgements: do we want him as an author?
Edited the beginning of the related-work section (intro paragraphs and strategy section)

\subsection{RTS AI Decomposition}
\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{figures/levels_abstraction.pdf}
    \caption{RTS AI levels of abstraction and their properties: uncertainty (from intentions and from partial observation) is going higher as the abstraction levels are raising. The timings on the right correspond to an estimate of the duration of a behavior switch in StarCraft. Spatial and temporal reasoning are indicated for part for which greedy solutions are really not efficient.}
    \label{fig:levels-abstraction}
\end{figure}
As work on RTS AI has to address several of the aforementioned problems together, in this section, we divide the existing work on RTS AI in three different levels of abstraction, widely known by RTS players: strategy, tactics and units control (micro-management), see Figure~\ref{fig:levels-abstraction}. We consider strategy to be everything related to the \textit{technology tree}, buildings, upgrades, and army composition. It is the most deliberative level, as a player performs a strategy with future stances (aggressive, defensive, economy, technology) and tactics in mind. To change its strategy, we consider that a player need about 3 minutes: they should need to construct a few new buildings and/or research some upgrades and/or produce units changing the army composition. Tactics represent the attacks and defenses between groups of units. For the human player, decisions on tactics are made from spatial (exploiting the terrain) and temporal (army movements) reasoning, constrained on the possible types of attacks by the army composition of the player and their opponent. Unit control, also called micro-management, is how the player takes action to maximize the efficiency of its individual units (applying the tactics) in real-time (simultaneous) and adversary conditions. As shown in Figure~\ref{fig:levels-abstraction}, strategy constrains possible tactics which in turn constrain units control. On a prediction/adaptation point of view, the units observation level gives information about the opponent's tactics and strategy. Flaws in opponents tactics can also steer our strategy into exploiting them.\\
{\color{blue}
perhaps TODO examples?
}

@@

In commercial RTS games, strategies are often ``hard coded'' with finite state machines (FSM) \cite{FSM_AIGameProgWisdom2003}. The idea behind FSMs is to decompose an object's behavior into easily manageable states, such as ``attacking'', ``gathering resources'' or ``repairing'' and establish the conditions that trigger transitions between them. Hierarchical FSM give higher (strategic) states some control over lower ones (actions). However, strategic decisions have to be made in a context, not in a vacuum. It is hard to encode dynamic, adaptive behaviors in (H)FSM due to the exponential number of transitions (between states) to account for. To this end, several approaches have been developed like hierarchical task networks, behavior trees and the comeback of STRIPS \cite{FikesSTRIPS} planning in the industry \cite{orkinGDC_FEAR}. Regardless, one cannot but notice that adaptability is not the strong point of industrial RTS AI. In the context of RTS AI strategy adaptation is related to the idea of opponent modeling. An adapting AI needs to keep track of what the enemy is doing in order to estimate the probability that they will use a specific strategy and adapt its own strategy accordingly.

In academic RTS AI research, the main methods used for strategic decision-making have been case-based reasoning (CBR) and planning or plan recognition \cite{LTW,CBR_Planning,OntanonCBR,HTNPlanning,Ramirez}. Aha et al. \cite{LTW} used CBR to perform dynamic plan retrieval extracted from domain knowledge in Wargus (Warcraft II clone). Onta\~{n}\'{o}n et al. \cite{CBR_Planning} base real-time case-based planning (CBP) system on a plan dependency graph which is learned from human demonstration. In \cite{OntanonCBR,PlanRetrieval}, they use CBR learned from expert demonstrations on Wargus. They update their previous CBP approach by using situation assessment for retrieving the ``right'' plan, and improve the speed of CBP by using a decision tree to select relevant features. Hsieh and Sun \cite{HsiehS08} based their work on \cite{LTW} CBR model and used StarCraft replays to construct states and building sequences (``build orders''). Schadd et al. \cite{SchaddBS07} describe opponent modeling through hierarchically structured models of the opponent behavior and they applied their work to the Spring RTS (Total Annihilation clone). Hoang et al. \cite{HTNPlanning} use hierarchical task networks (HTN) to model strategies in a first person shooter with the goal to use HTN planners. %Madeira et al. \cite{Madeira06}

Weber and Mateas \cite{WeberCig09} proposed ``a data mining approach to strategy prediction'' and performed supervised learning (from buildings features) on labeled StarCraft replays. Dereszynski et al. \cite{HMMstrat_RTS_AIIDE11} used HMM to learn the transition probabilities of sequences of constructions and kept the most probables to produce probabilistic behavior models (in StarCraft). Synnaeve and Bessi\`{e}re \cite{SynnaeveOpeningCig11} extended \cite{WeberCig09} and used a Bayesian semi-supervised (labeling based on EM clustering considering appropriate features) model to learn from replays and predict openings (early game strategies) from StarCraft replays. Then \cite{SynnaeveAIIDE11}, they presented an unsupervised learning Bayesian model for tech-tree prediction, still using replays.

One final consideration is that RTS games are typically partially observable. For example games like StarCraft implement the ``fog of war'' idea, which basically means that a player can only see the areas of the map close to her own units. Areas of the map away from the field of view of individual units are not observable. Therefore, in order to play an RTS game, players need to scout the opponent in order to obtain information about the opponent's strategy. Most commercial games avoid this problem by using perfect information all the time (cheating). In order to make the human player believe the built-in AI of this games does scouting, sometimes they simulate some scouting tasks as Bob Fitch described in his AIIDE 2011 keynote for the WarCraft game series and StarCraft: Broodwar. %{\color{blue}XXX (perhaps move it above, to the start of subsection}

@@

\section*{Acknowledgments} {\color{blue} This research is partially funded by projects ... and ... }

>>>>

%\subsection{RTS AI Decomposition}
\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{figures/levels_abstraction.pdf}
    \caption{RTS AI levels of abstraction and their properties: uncertainty (from intentions and from partial observation) is going higher as the abstraction levels are raising. The timings on the right correspond to an estimate of the duration of a behavior switch in StarCraft. Spatial and temporal reasoning are indicated for part for which greedy solutions are really not efficient.}
    \label{fig:levels-abstraction}
\end{figure}

AI systems that play RTS games need to address most, if not all, the aforementioned problems together. Therefore, in order to classify existing work on RTS AI, we will divide the existing work according to three levels of abstraction (division that is widely used by RTS players): strategy, tactics and unit control (micro-management).

Figure~\ref{fig:levels-abstraction} graphically illustrates how strategy, tactics and unit control are three points in a continuum scale where strategy corresponds to decisions making processes that affect long spans of time (several minutes in the case of Starcraft), unit control corresponds to low-level reactive second-by-second individual unit control, and tactics sit in the middle. Also, strategic decisions reason about the whole game at once, whereas tactical decisions are localized, and affect only specific groups of units; finally unit control techniques control only individual units. Typically, strategic decisions constraint future tactical decisions, which in turn constrain unit control. Moreover, information gathered while performing unit control, can cause reconsideration of the tactics being employed; which could trigger further strategical reasoning.

Following this idea, we consider strategy to be everything related to the technology trees, build-order\footnote{The {\em build-order} is the specific sequence in which buildings of different types will be constructed at the beginning of a game, and completely determines the long-term strategy of a player.}, upgrades, and army composition. It is the most deliberative level, as a player selected and performs a strategy with future stances (aggressive, defensive, economy, technology) and tactics in mind. We consider tactics, everything related to confrontations between groups of units. Tactical reasoning involves both spatial (exploiting the terrain) and temporal (army movements) reasoning, constrained on the possible types of attacks by the army composition of the player and their opponent. Finally, we will call unit control (also known as micro-management) to how the player controls individual units to maximize their efficiency (when executing tactics) in real-time and adversary conditions.

% Example:
For example, after starting a game, a player might decide to use a {\em rushing} strategy (which involves quickly building an army and sending it to attack as early as possible in the game); then, when performing the attack use a {\em surrounding} tactic, where the player tries to surround the enemy cutting potential escape routes; finally, while executing the surrounding tactic, the player might decide to use unit control techniques that command individual units to perform repeated {\em attack and flee} movements, to maximize the efficiency of each of the units being used in the attack.

@@

Strategic decision making is a hard problem, which has been addressed using many AI paradigms. Most of the work in this area falls in one of three categories: hard-coded approaches, planning-based approaches, and machine learning-based approaches. Let us cover each of these categories in turn.

Hard-coded approaches have been extensively used in commercial RTS games. The most common approaches make use finite state machines (FSM) \cite{FSM_AIGameProgWisdom2003} in order to let the AI author hard-code the strategy that the AI will employ. The idea behind FSMs is to decompose the AI behavior into easily manageable states, such as ``attacking'', ``gathering resources'' or ``repairing'' and establish the conditions that trigger transitions between them. Commercial approaches also include Hierarchical FSMs, in which FSMS are composed hierarchically. These hard-coded approaches have achieved a significant amount of success, and, as we will discuss later, have also been used in many academic RTS AI research systems, as discussed in Section \ref{sec:bot}. However, strategic decision making is a hard problem, and these hard-coded approaches struggle to encode dynamic, adaptive behaviors.

% XXX: Santi: I've commented out this paragraph, since it has nothing to do with RTS games. There is no commercial RTS game using planning (neither STRIPS nor HTN) that I'm aware of. The references below refer to FEAR, which is an FPS game.
%To this end, several approaches have been developed like hierarchical task networks, behavior trees and the comeback of STRIPS \cite{FikesSTRIPS} planning in the industry \cite{orkinGDC_FEAR}. Regardless, one cannot but notice that adaptability is not the strong point of industrial RTS AI. In the context of RTS AI strategy adaptation is related to the idea of opponent modeling. An adapting AI needs to keep track of what the enemy is doing in order to estimate the probability that they will use a specific strategy and adapt its own strategy accordingly.

Approaches using planning techniques have also been explored in the literature. For example Onta\~{n}\'{o}n et al. \cite{CBR_Planning} explored the use of real-time case-based planning (CBP) in the domain of Wargus (a Warcraft II clone). In their work, they used human demonstration to learn plans, that are then composed at run-time in order to form full-fledges strategies to play the game. In \cite{PlanRetrieval} they their previous CBP approach by using situation assessment for retrieving the ``right'' plan, and improve the speed of CBP by using a decision tree to select relevant features. Hierarchical Task-Network (HTN) planning has also been explored with some success in the context of simpler first-person shooter games. Planning approaches offer more adaptivity of the AI strategy compared to hard-coded approaches. However, the real-time constraints of RTS games limit the planning approaches that can be applied, being HTN and case-based planning the only ones explored so far.

Concerning machine learning-based approaches, Weber and Mateas \cite{WeberCig09} proposed ``a data mining approach to strategy prediction'' and performed supervised learning (from buildings features) on labeled StarCraft replays. Dereszynski et al. \cite{HMMstrat_RTS_AIIDE11} used HMM to learn the transition probabilities of sequences of constructions and kept the most probables to produce probabilistic behavior models (in StarCraft). Synnaeve and Bessi\`{e}re \cite{SynnaeveOpeningCig11} extended \cite{WeberCig09} and used a Bayesian semi-supervised (labeling based on EM clustering considering appropriate features) model to learn from replays and predict openings (early game strategies) from StarCraft replays. Then \cite{SynnaeveAIIDE11}, they presented an unsupervised learning Bayesian model for tech-tree prediction, still using replays.

Also falling into the machine-learning category, a significant group of researchers has explored case-based reasoning (CBR) \cite{Aamodt94CBR} approaches for strategic decision making. For example Aha et al. \cite{LTW} used CBR to perform dynamic plan retrieval in the Wargus domain. Hsieh and Sun \cite{HsiehS08} based their work on \cite{LTW} CBR model and used StarCraft replays to construct states and building sequences (``build orders''). Finally, Schadd et al. \cite{SchaddBS07} applied a CBR approach to opponent modeling through hierarchically structured models of the opponent behavior and they applied their work to the Spring RTS (Total Annihilation clone).

One final consideration is that RTS games are typically partially observable. For example games like StarCraft implement the ``fog of war'' idea, which basically means that a player can only see the areas of the map close to her own units. Areas of the map away from the field of view of individual units are not observable. Therefore, in order to play an RTS game, players need to scout the opponent in order to obtain information about the opponent's strategy. Very few of the previous approaches deal with this problem, and use perfect information all the time (in the case of commercial games, most AI implementations cheat, since they have perfect information). In commercial games, in order to make the human player believe the built-in AI of this games does scouting, sometimes they simulate some scouting tasks as Bob Fitch described in his AIIDE 2011 keynote for the WarCraft game series and StarCraft: Broodwar. %{\color{blue}XXX (perhaps move it above, to the start of subsection}

@@

\section*{Acknowledgments} {\color{blue} This research is partially funded by projects ... and ... . The authors would like to thank Florian Richoux, author of the AIUR bot, for his helpful comments and help in early drafts of this paper. }
